---
title: "Cod-Flounder eco-evo model"
date: "05/21/2024"
output: 
  html_document:
      code_folding: show
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "left", fig.width = 6, fig.height = 3, out.width = "67%")
library(tidyverse)
library(deSolve)
library(cowplot)

theme_set(theme_bw())
theme_update(panel.grid = element_blank(),
             strip.background = element_rect(fill = "white"))
```

## Main functions

As per Eqs. 26-27.

```{r main_funcs}
erf <- function(x) 2 * pnorm(x * sqrt(2)) - 1

dNdt <- function(m, s, eta, theta, rho, N, alpha0, mF, sF, Fpop, w, zstar, tau, Z, nu, k) {
  N * (rho / theta^2 * (theta^2 - (zstar - m)^2 - s^2) - 
         eta / 2 * (erf(m / sqrt(2 * s^2 + tau^2)) + 1) - 
         k / 2* (erf((m - Z) / sqrt(2 * s^2 + nu^2))) - 
         N * alpha0 * sqrt(w^2/(4 * s^2 + w^2)) - 
         Fpop * (alpha0 * sqrt(w^2 / (2 * s^2 + 2 * sF^2 + w^2)) * 
                exp(-(m - mF)^2 / (2 * s^2 + 2 * sF^2 + w^2)))
       )
}


dmdt <- function(m, s, eta, theta, rho, alpha0, mF, sF, Fpop, w, zstar, tau, Z, nu, k, h2) {
  h2 * ( (2 * rho * s^2 / theta^2) * (zstar - m) - 
           eta * s^2 / sqrt(pi * (2 * s^2 + tau^2)) * exp(-m^2 / (2 * s^2 + tau^2)) - 
           k * s^2 / sqrt(pi * (2*s^2 + nu^2)) * exp(-(m - Z)^2 / (2*s^2 + nu^2)) -
           Fpop * (2 * s^2 * w * (mF - m) * alpha0 / (2 * s^2 + 2 * sF^2 + w^2)^(3/2) * 
                  exp(-(m - mF)^2 / (2 * s^2 + 2 * sF^2 + w^2))) 
         )
}

```

# Fig. 1

## Density examples

Illustrating body size distributions before and after fishing.

```{r DensityExample}

equildf1 <- tibble(`body size` = c(0,1.4), density = c(1,1), species = c("flounder", "cod"), txt = c("b", "c"))
gg_equil1 <- tibble(`body size` = seq(-2.4, 2.4, l = 201)) %>%
  mutate(flounder = exp(-(`body size`)^2 / (2 * 0.55^4)),
         cod = 1 * exp(-(`body size` - 1.4)^2 / (2 * 0.55^4))) %>%
  pivot_longer(cols = c(flounder, cod), names_to = "species", values_to = "density") %>%
  mutate(density = if_else(density < 1e-4, NA_real_, density)) %>%
  ggplot(aes(x = `body size`, y = density, ymin = 0, ymax = density,
             colour = species, fill = species)) +
  geom_line(na.rm = TRUE) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_shape_manual(values = c(22,21)) + 
  geom_ribbon(colour = NA, alpha = 0.2, na.rm = TRUE) +
  scale_colour_manual(values = c("cornflowerblue", "darkseagreen")) + #"darkseagreen" for manuscript #"goldenrod" for presentations
  scale_fill_manual(values = c("cornflowerblue", "darkseagreen")) +
  scale_y_continuous(limits = c(0,1.25), breaks = c(0, .25, .5, .75, 1)) +
  geom_point(data = equildf1, aes(x = `body size`, y = density, shape = species, color = species), size = 3, color = "black") +
  geom_text(data = equildf1, aes(x = `body size`, y = density, label = txt), nudge_x = 0.2, nudge_y = 0.08, color = "black") +
  labs(x = expression(paste("mean body size ", mu))) + 
  theme(#legend.position = "bottom", legend.direction = "horizontal",
        #legend.box.margin = margin(t = -10), 
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

equildf2 <- tibble(`body size` = c(0,-1.0), density = c(1,1), species = c("flounder", "cod"), txt = c("b", "a"))
gg_equil2 <- tibble(`body size` = seq(-2.4, 2.4, l = 201)) %>%
  mutate(flounder = exp(-(`body size`)^2 / (2 * 0.55^4)),
         cod = 1 * exp(-(`body size` + 1.0)^2 / (2 * 0.55^4))) %>%
  pivot_longer(cols = c(flounder, cod), names_to = "species", values_to = "density") %>%
  mutate(density = if_else(density < 1e-4, NA_real_, density)) %>%
  ggplot(aes(x = `body size`, y = density, ymin = 0, ymax = density,
             colour = species, fill = species)) +
  geom_line(na.rm = TRUE) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_shape_manual(values = c(23,21)) + 
  geom_ribbon(colour = NA, alpha = 0.2, na.rm = TRUE) +
  scale_colour_manual(values = c("cornflowerblue", "darkseagreen")) +
  scale_fill_manual(values = c("cornflowerblue", "darkseagreen")) +
  scale_y_continuous(limits = c(0,1.25), breaks = c(0, .25, .5, .75, 1)) +
  geom_point(data = equildf2, aes(x = `body size`, y = density, shape = species, color = species), size = 3, color = "black") +
  geom_text(data = equildf2, aes(x = `body size`, y = density, label = txt), nudge_x = 0.2, nudge_y = 0.08, color = "black") +
  labs(x = expression(paste("mean body size ", mu))) +
  theme(#legend.position = "none", legend.direction = "horizontal",
        #legend.box.margin = margin(t = -10), 
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) 

```


## Trait change equilibria

Illustrating the body size dependent rate of change for cod body size, with and without fishing.

```{r bodysize_nofishing}
equildf3 <- tibble(x = c(-1.4, 0, 1.4), y = c(0,0,0), sh = factor(c("small cod","flounder","large cod"), levels = c("small cod","flounder","large cod")), txt = c("a", "b", "c"))
bodysize_eq_nofish <- tibble(m = seq(-2.0, 2.0, l = 2001)) %>%
  mutate(nofishing = dmdt(m, mF = 0, s = 1/2, theta = 4, eta = 0, alpha0 = 1 / sqrt(2), 
                        rho = 12, sF = 1/2, Fpop = 8, w = 1, h2 = 1.0, zstar = 0, 
                        k = 0, tau = 1/2, Z = 1, nu = 3/2)) %>%
  pivot_longer(cols = contains("fishing")) %>%
  mutate(name = fct_relevel(name, "nofishing")) %>%
  filter(name == "nofishing") %>%  # & value > -0.001 & value < 0.001 ## to check x-axis intersects
  ggplot(aes(x = m, y = value)) +
  geom_line(colour = "cornflowerblue") +
  scale_x_continuous(expand = expansion(0,0)) +
  geom_hline(yintercept = 0, alpha = 0.4, linetype = "solid") +
  geom_vline(xintercept = 1, alpha = 1, linetype = "dashed") +
  geom_point(equildf3, mapping=aes(x,y, shape = (sh), fill = (sh)), size  = 3) +
  scale_shape_manual(values = c(23,21,22)) +
  scale_color_manual(values =  c("cornflowerblue", "darkseagreen", "cornflowerblue")) +
  scale_fill_manual(values =  c("cornflowerblue", "darkseagreen", "cornflowerblue")) +
  geom_text(data = equildf3, aes(x, y, label = txt), nudge_x = 0.25, nudge_y = 0.08, color = "black") +
  labs(x = expression(paste("mean body size ", mu)), y = expression(paste(d~mu/d~t)), shape = "", fill = "") +
  scale_y_continuous(limits = c(-0.6, 0.6)) +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

# Extract legend
cowlegend <- get_legend(bodysize_eq_nofish + 
                          theme(legend.direction = "horizontal"))


equildf4 <- tibble(x = c(-1.46, 0), y = c(0, 0), sh = c(1, 2), txt = c("a", "b"))
bodysize_eq_fish <- tibble(m = seq(-2.0, 2.0, l = 2001)) %>%
mutate(nofishing = dmdt(m, mF = 0, s = 1/2, theta = 4, eta = 0, alpha0 = 1 / sqrt(2), 
                        rho = 12, sF = 1/2, Fpop = 8, w = 1, h2 = 1.0, zstar = 0, 
                        k = 5, tau = 1/2, Z = 1, nu = 3/2)) %>%
  pivot_longer(cols = contains("fishing")) %>%
  mutate(name = fct_relevel(name, "nofishing")) %>%
  filter(name == "nofishing") %>% 
  ggplot(aes(x = m, y = value)) +
  geom_line(colour = "cornflowerblue") +
  scale_shape_manual(values = c(23, 21), guide = "none") +
  geom_hline(yintercept = 0, alpha = 0.4, linetype = "solid") +
  geom_vline(xintercept = 1, alpha = 1, linetype = "dashed") +
  scale_x_continuous(expand = expansion(0,0)) +
  geom_point(equildf4, mapping=aes(x,y, shape = as.factor(sh)), size  = 3,fill = c("cornflowerblue", "darkseagreen")) +
  geom_text(data = equildf4, aes(x, y, label = txt), nudge_x = 0.25, nudge_y = 0.08, color = "black") +
  labs(x = expression(paste("mean body size ", mu)), y = expression(paste(d~mu/d~t)), shape = "species") +
  scale_y_continuous(limits = c(-0.6, 0.6)) +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

## Fig. 1 Output

```{r figure1, out.width="80%", fig.width = 6.5, fig.height = 7.8}
sizegrid <- cowplot::plot_grid(gg_equil1 + theme(legend.position = "none"), 
                                  bodysize_eq_nofish + theme(legend.position = "none"), 
                               gg_equil1 + theme(legend.position = "none"), 
                                  bodysize_eq_fish + theme(legend.position = "none"), 
                               gg_equil2 + theme(legend.position = "none"), 
                                  bodysize_eq_nofish + theme(legend.position = "none"),
                               NULL, NULL,
                   ncol = 2,
                   rel_heights = c(1,1,1,0.15),
                   rel_widths = c(1,0.8), labels = c("A", "B", "C", "D", "E", "F"))
sizegrid + draw_grob(cowlegend, x = 0, y = 0, width = 1, height = 0.2/3.2)

ggsave(filename = "../fig/fig1.pdf", plot = sizegrid + draw_grob(cowlegend, x = 0, y = 0, width = 1, height = 0.2/3.2), width = 5*1.3, height = 6*1.3)
```

# Full model

## Parameterization 

Parameters with their respective code variable names in parenthesis, and their default
values.

| Parameter         | Value    | Explanation                                      |
|------------------|------------------|-------------------------------------|
| $z$ (z)           | variable | Cod body size                                    |
| $\mu$ (m)         | variable | Cod mean body size                               |
| $\mu_{F}$ (mF)    | 0        | Flounder mean body size                          |
| $\sigma$ (s)      | $1/2$    | Cod body size standard deviation                 |
| $\sigma_{F}$ (sF) | $1/2$    | Flounder body size standard deviation            |
| $\theta$ (theta)  | 5        | Trait range width where growth is still possible |
| $\varrho$ (rho)   | 5       | Growth magnitude parameter                       |
| $F$ (Fpop)        | 4      | Flounder competition magnitude                   |
| $\omega$ (w)      | 1/2       | Trait distance competition parameter             |
| $h^{2}$ (h2)      | 1/2      | Heritability                                     |
| $z^{*}$ (zstar)   | 1        | Cod ideal growth body size                       |
| $\eta$ (eta)      | 0-2      | Fishing intensity                                |
| $\tau$ (tau)      | $1/2$    | Fishing intensity transition                     |
| $Z$ (Z)           | $1$      | Hypoxia intensity threshold                      |
| $\nu$ (nu)        | $3/2$    | Hypoxia intensity transition                     |
| $\kappa$ (k)      | 2        | Hypoxic effect magnitude                         |

## ODE functions

```{r ode_eq}
eqs <- function(time, state, p) {
  N <- state[1]
  m <- state[2]
  Ndot <- dNdt(m, p$s, p$eta, p$theta, p$rho, N, p$alpha0, p$mF, p$sF, p$Fpop, p$w, p$zstar, p$tau, p$Z, p$nu, p$k)
  mdot <- dmdt(m, p$s, p$eta, p$theta, p$rho, p$alpha0, p$mF, p$sF, p$Fpop, p$w, p$zstar, p$tau, p$Z, p$nu, p$k, p$h2)
  list(c(Ndot, mdot))
}
```

### Fig. 2 Parameterized model fishing scenarios

```{r bodysize_fishing, out.width = "80%"}

 gg_mu_fishing <- tibble(m = seq(-2.0, 2.0, l = 201)) %>% 
  # eta fishing intensity parameter
  mutate(nofishing = dmdt(m, mF = 0, s = 1/2, theta = 5, eta = 0, alpha0 = 1 / sqrt(2), 
                        rho = 5, sF = 1/2, Fpop = 4, w = 1/2, h2 = 0.5, zstar = 1, 
                        k = 2, tau = 1/2, Z = 1, nu = 3/2),
       fishing = dmdt(m, mF = 0, s = 1/2, theta = 5, eta = 1, alpha0 = 1 / sqrt(2), 
                        rho = 5, sF = 1/2, Fpop = 4, w = 1/2, h2 = 0.5, zstar = 1, 
                        k = 2, tau = 1/2, Z = 1, nu = 3/2),
       intfishing = dmdt(m, mF = 0, s = 1/2, theta = 5, eta = 2, alpha0 = 1 / sqrt(2), 
                        rho = 5, sF = 1/2, Fpop = 4, w = 1/2, h2 = 0.5, zstar = 1, 
                        k = 2, tau = 1/2, Z = 1, nu = 3/2)) %>%
  pivot_longer(cols = contains("fishing")) %>%
  mutate(name = fct_relevel(name, "nofishing", "fishing", "intfishing")) %>%
  mutate(name = recode(name, "intfishing" = "bold(C)~-~Intense~fishing",
                       "fishing" = "bold(B)~-~Fishing",
                       "nofishing" = "bold(A)~-~No~fishing")) %>%
  ggplot(aes(x = m, y = value)) +
  geom_line(colour = "steelblue") +
  geom_hline(yintercept = 0, alpha = 0.4, linetype = "dashed") +
  labs(x = expression(paste("Body size (", mu, ")")), y = expression(paste("Body size rate of change (", d~mu/d~t, ")"))) +
  scale_y_continuous(limits = c(-0.2, 0.2)) +
  facet_grid(. ~ name, labeller = label_parsed)
gg_mu_fishing
ggsave(plot = gg_mu_fishing, filename = "../fig/fig2.pdf", width = 7.0, height = 2.7)
```

