---
title: "Cod-Flounder eco-evo model"
author: "Ohlsson et al."
date: "08/21/2023"
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "left", fig.width = 6, fig.height = 3, out.width = "40%")
library(tidyverse)
library(deSolve)
library(cowplot)

theme_set(theme_bw())
theme_update(panel.grid = element_blank(),
             strip.background = element_rect(fill = "white"))
```

# Eco-evolutionary model

Parameters with their respective variable names in parenthesis, default values.

| Parameter    | Value    | Explanation                                      |
|--------------|----------|--------------------------------------------------|
| $z$ (z)         | variable | Cod body size                                    |
| $\mu$ (m)       | variable | Cod mean body size                               |
| $\mu_{F}$ (mF)    | 0        | Flounder mean body size                          |
| $\sigma$ (s)     | $1/2$    | Cod body size standard deviation                 |
| $\sigma_{F}$ (sF) | $1/3$    | Flounder body size standard deviation            |
| $\theta$ (theta)      | 5        | Trait range width where growth is still possible |
| $\varrho$ (rho)    | 15       | Growth magnitude parameter                       |
| $F$ (Fpop)         | 10       | Flounder competition magnitude                   |
| $\omega$ (w)     | 1        | Trait distance competition parameter             |
| $h^{2}$ (h2)     | 0.1      | Heritability                                     |
| $z^{*}$ (zstar)     | 1        | Cod ideal growth body size                       |
| $\eta$ (eta)       | 0, 1, 2  | Fishing intensity                                |
| $\kappa$ (o)    | 3        | Hypoxic effect magnitude                         |

## Helper functions

As per supplementary functions S12-S17.

```{r helper_funcs}
#Defining complementary error function
erf <- function(x) 2 * pnorm(x * sqrt(2)) - 1
erfc <- function(x) 2 * pnorm(x * sqrt(2), lower = FALSE)

b <- function(zstar, m, s, theta, rho, o) -rho * (-theta^2 + (zstar - m)^2 + s^2) / theta^2 - 1/2 * o * erfc((zstar-m) / (sqrt(2)*s)) 
u <- function(m, s, eta) eta * exp(m + s^2 / 2)
falpha <- function(m, mF, s, sF, w) {
  W <- 2 * (s^2 + sF^2) + w^2
  exp(-(m - mF)^2 / W) * w / sqrt(W)
}

g <- function(zstar, m, s, theta, rho, o) -exp(-(zstar-m)^2 / (2*s^2)) * o * s / sqrt(2*pi) + 2 * (zstar - m) * rho * s^2 / theta^2 
  
v <- function(m, s, eta)  s^2 * eta * exp(m + s^2 / 2)
fbeta <- function(m, mF, s, sF, w) {
  falpha(m, mF, s, sF, w) * 2 * s^2 * (mF - m) / (2 * (s^2 + sF^2) + w^2)
}
```

## Main functions

As per Supplementary functions S6-S7.

```{r main_funcs}
dNdt <- function(m, s, eta, theta, rho, N, alpha0, mF, sF, Fpop, w, zstar, o) {
  N * (b(zstar, m, s, theta, rho, o) - u(m, s, eta) - N * alpha0 - Fpop * falpha(m, mF, s, sF, w))
}


dmdt <- function(m, s, theta, rho, eta, Fpop, mF, sF, w, h2, zstar, o) {
  h2 * (g(zstar, m, s, theta, rho, o) - v(m, s, eta) - Fpop * fbeta(m, mF, s, sF, w))
}
```


# Plotting

## Density examples

Illustrating body size distributions before and after fishing.

```{r DensityExample}

equildf1 <- tibble(`body size` = c(0,1.4), density = c(1,1), species = c("flounder", "cod"), txt = c("b", "c"))
gg_equil1 <- tibble(`body size` = seq(-2.4, 2.4, l = 201)) %>%
  mutate(flounder = exp(-(`body size`)^2 / (2 * 0.55^4)),
         cod = 1 * exp(-(`body size` - 1.4)^2 / (2 * 0.55^4))) %>%
  pivot_longer(cols = c(flounder, cod), names_to = "species", values_to = "density") %>%
  mutate(density = if_else(density < 1e-4, NA_real_, density)) %>%
  ggplot(aes(x = `body size`, y = density, ymin = 0, ymax = density,
             colour = species, fill = species)) +
  geom_line(na.rm = TRUE) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_shape_manual(values = c(22,21)) + 
  geom_ribbon(colour = NA, alpha = 0.2, na.rm = TRUE) +
  scale_colour_manual(values = c("cornflowerblue", "darkseagreen")) + #"darkseagreen" for manuscript #"goldenrod" for presentations
  scale_fill_manual(values = c("cornflowerblue", "darkseagreen")) +
  scale_y_continuous(limits = c(0,1.25), breaks = c(0, .25, .5, .75, 1)) +
  geom_point(data = equildf1, aes(x = `body size`, y = density, shape = species, color = species), size = 3, color = "black") +
  geom_text(data = equildf1, aes(x = `body size`, y = density, label = txt), nudge_x = 0.2, nudge_y = 0.08, color = "black") +
  labs(x = expression(paste("mean body size ", mu))) + 
  theme(#legend.position = "bottom", legend.direction = "horizontal",
        #legend.box.margin = margin(t = -10), 
        panel.grid = element_blank())
#gg_equil1

equildf2 <- tibble(`body size` = c(0,-1.0), density = c(1,1), species = c("flounder", "cod"), txt = c("b", "a"))
gg_equil2 <- tibble(`body size` = seq(-2.4, 2.4, l = 201)) %>%
  mutate(flounder = exp(-(`body size`)^2 / (2 * 0.55^4)),
         cod = 1 * exp(-(`body size` + 1.0)^2 / (2 * 0.55^4))) %>%
  pivot_longer(cols = c(flounder, cod), names_to = "species", values_to = "density") %>%
  mutate(density = if_else(density < 1e-4, NA_real_, density)) %>%
  ggplot(aes(x = `body size`, y = density, ymin = 0, ymax = density,
             colour = species, fill = species)) +
  geom_line(na.rm = TRUE) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_shape_manual(values = c(23,21)) + 
  geom_ribbon(colour = NA, alpha = 0.2, na.rm = TRUE) +
  scale_colour_manual(values = c("cornflowerblue", "darkseagreen")) +
  scale_fill_manual(values = c("cornflowerblue", "darkseagreen")) +
  scale_y_continuous(limits = c(0,1.25), breaks = c(0, .25, .5, .75, 1)) +
  geom_point(data = equildf2, aes(x = `body size`, y = density, shape = species, color = species), size = 3, color = "black") +
  geom_text(data = equildf2, aes(x = `body size`, y = density, label = txt), nudge_x = 0.2, nudge_y = 0.08, color = "black") +
  labs(x = expression(paste("mean body size ", mu))) +
  theme(#legend.position = "none", legend.direction = "horizontal",
        #legend.box.margin = margin(t = -10), 
        panel.grid = element_blank()) 
#gg_equil2
  
```


## Trait change equilibria

Illustrating the body size dependent rate of change for cod body size, with and without fishing.

```{r bodysize_nofishing}
equildf3 <- tibble(x = c(-1.22, 0, 1.23), y = c(0,0,0), sh = factor(c("small cod","flounder","large cod"), levels = c("small cod","flounder","large cod")), txt = c("a", "b", "c"))
bodysize_eq_nofish <- tibble(m = seq(-2.5, 2.5, l = 201)) %>%
mutate(nofishing = dmdt(m, mF = 0, s = 1/2, theta = 4, eta = 0, rho = 20, sF = 1/3, Fpop = 6.8, w = 1, h2 = 1, zstar = 0, o = 0)) %>%
  pivot_longer(cols = contains("fishing")) %>%
  mutate(name = fct_relevel(name, "nofishing")) %>%
  filter(name == "nofishing") %>% 
  ggplot(aes(x = m, y = value)) +
  geom_line(colour = "cornflowerblue") +
  scale_x_continuous(expand = expansion(0,0)) +
  geom_hline(yintercept = 0, alpha = 0.4, linetype = "solid") +
  geom_vline(xintercept = 1, alpha = 1, linetype = "dashed") +
  geom_point(equildf3, mapping=aes(x,y, shape = (sh), fill = (sh)), size  = 3) +
  scale_shape_manual(values = c(23,21,22)) +
  scale_color_manual(values =  c("cornflowerblue", "darkseagreen", "cornflowerblue")) +
  scale_fill_manual(values =  c("cornflowerblue", "darkseagreen", "cornflowerblue")) +
  geom_text(data = equildf3, aes(x, y, label = txt), nudge_x = 0.25, nudge_y = 0.08, color = "black") +
  labs(x = expression(paste("mean body size ", mu)), y = expression(paste(d~mu/d~t)), shape = "", fill = "") +
  scale_y_continuous(limits = c(-0.75, 0.75)) +
  theme(panel.grid = element_blank())
#bodysize_eq_nofish
#ggsave(plot = gg_mu_nofishing, filename = "../fig/mu_nofishing.png", width = 4, height =4)  
cowlegend <- get_legend(bodysize_eq_nofish + 
                          theme(legend.direction = "horizontal"))


equildf4 <- tibble(x = c(-1.3, 0), y = c(0, 0), sh = c(1, 2), txt = c("a", "b"))
bodysize_eq_fish <- tibble(m = seq(-2.5, 2.5, l = 201)) %>%
mutate(nofishing = dmdt(m, mF = 0, s = 1/2, theta = 4, eta = 1, rho = 20, sF = 1/3, Fpop = 6.8, w = 1, h2 = 1, zstar = 0, o = 0)) %>%
  pivot_longer(cols = contains("fishing")) %>%
  mutate(name = fct_relevel(name, "nofishing")) %>%
  filter(name == "nofishing") %>% 
  ggplot(aes(x = m, y = value)) +
  geom_line(colour = "cornflowerblue") +
  scale_shape_manual(values = c(23, 21), guide = "none") +
  geom_hline(yintercept = 0, alpha = 0.4, linetype = "solid") +
  geom_vline(xintercept = 1, alpha = 1, linetype = "dashed") +
  scale_x_continuous(expand = expansion(0,0)) +
  geom_point(equildf4, mapping=aes(x,y, shape = as.factor(sh)), size  = 3,fill = c("cornflowerblue", "darkseagreen")) +
  geom_text(data = equildf4, aes(x, y, label = txt), nudge_x = 0.25, nudge_y = 0.08, color = "black") +
  labs(x = expression(paste("mean body size ", mu)), y = expression(paste(d~mu/d~t)), shape = "species") +
  scale_y_continuous(limits = c(-0.75, 0.75)) +
  theme(panel.grid = element_blank())
#bodysize_eq_fish
```

### Compound plot

```{r figure1, out.width="100%", fig.width = 6.5, fig.height = 7.8}
sizegrid <- cowplot::plot_grid(gg_equil1 + theme(legend.position = "none"), 
                                  bodysize_eq_nofish + theme(legend.position = "none"), 
                               gg_equil1 + theme(legend.position = "none"), 
                                  bodysize_eq_fish + theme(legend.position = "none"), 
                               gg_equil2 + theme(legend.position = "none"), 
                                  bodysize_eq_nofish + theme(legend.position = "none"),
                               NULL, NULL,
                   ncol = 2,
                   rel_heights = c(1,1,1,0.15),
                   rel_widths = c(1,0.8), labels = c("A", "B", "C", "D", "E", "F"))
sizegrid + draw_grob(cowlegend, x = 0, y = 0, width = 1, height = 0.2/3.2)

#ggsave(filename = "../fig/sizegrid_greengold.pdf", plot = sizegrid + draw_grob(cowlegend, x = 0, y = 0, width = 1, height = 0.2/3.2), width = 5*1.3, height = 6*1.3)
```

# Parameterized model

```{r bodysize_fishing, out.width = "80%"}
 gg_mu_fishing <- tibble(m = seq(-2.5, 2.5, l = 201)) %>% 
  mutate(nofishing = dmdt(m, s = 1/2, theta = 5, rho = 15, eta = 0,
                          Fpop = 10, mF = 0, sF = 1/3, w = 1, 
                          h2 = 0.1, zstar = 1, o = 3),
       fishing = dmdt(m, s = 1/2, theta = 5, rho = 15, eta = 1,
                          Fpop = 10, mF = 0, sF = 1/3, w = 1, 
                      h2 = 0.1, zstar = 1, o = 3),
       intfishing = dmdt(m, s = 1/2, theta = 5, rho = 15, eta = 2,
                        Fpop = 10, mF = 0, sF = 1/3, w = 1, 
                        h2 = 0.1, zstar = 1, o = 3)) %>%
  pivot_longer(cols = contains("fishing")) %>%
  mutate(name = fct_relevel(name, "nofishing", "fishing", "intfishing")) %>%
  mutate(name = recode(name, "intfishing" = "Intense~fishing~(eta==2)",
                       "fishing" = "Fishing~(eta==1)",
                       "nofishing" = "No~fishing~(eta==0)")) %>%
  ggplot(aes(x = m, y = value)) +
  geom_line(colour = "steelblue") +
  geom_hline(yintercept = 0, alpha = 0.4, linetype = "dashed") +
  labs(x = expression(paste(mu)), y = expression(paste(d~mu/d~t))) +
  scale_y_continuous(limits = c(-0.15, 0.15)) +
  facet_grid(. ~ name, labeller = label_parsed)
gg_mu_fishing
#ggsave(plot = gg_mu_fishing, filename = "../fig/mu_scenarios.pdf", width = 5.5, height = 2.2)

```

# Bodysize development over time in different scenarios

With current parameterization, population size $N$ is only used to confirm that the population does not crash.

### ODE Equations

```{r ode_eq}
eqs <- function(time, state, p) {
  N <- state[1]
  m <- state[2]
  Ndot <- dNdt(m, p$s, p$eta, p$theta, p$rho, N, p$alpha0, p$mF, p$sF, p$Fpop, p$w, p$zstar, p$o)
  mdot <- dmdt(m, p$s, p$theta, p$rho, p$eta, p$Fpop, p$mF, p$sF, p$w, p$h2, p$zstar, p$o)
  list(c(Ndot, mdot))
}
```

### Starting with ideal cod body size, no fishing ($\eta = 0$) and hypoxia ($\omicron = 3$):

```{r healthy_with_nofishing_hypoxia, fig.width = 6, out.width = "67%"}
tibble(pars = list(list(mF = 0, s = 1/2, theta = 5, eta = 0, alpha0 = 1 / sqrt(2), 
                        rho = 15, sF = 1/3, Fpop = 10, w = 1, h2 = 0.1, zstar = 1, o = 3))) %>%
  mutate(ic = list(c(12, 1))) %>% 
  mutate(sol = map2(pars, ic, ~ode(.y, seq(0, 100, by = 0.01), eqs, .x) %>%
                      as.data.frame() %>%
                      as_tibble() %>%
                      rename(N = `1`, mu = `2`) %>%
                      pivot_longer(cols = c(N, mu)) %>%
                      mutate(name = fct_relevel(name, "N", "mu")))) %>%
  unnest(sol) %>% 
  #filter(name == "mu") %>%
  ggplot(aes(x = time, y = value)) +
  geom_line(colour = "steelblue") +
  facet_wrap(.~name, scales = "free_y")
```

### Starting with ideal cod body size, moderate fishing ($\eta = 1$) and hypoxia ($\omicron = 0$):

```{r healthy_with_fishing_hypoxia, fig.width = 6, out.width = "67%"}
tibble(pars = list(list(mF = 0, s = 1/2, theta = 5, eta = 1, alpha0 = 1 / sqrt(2), 
                        rho = 15, sF = 1/3, Fpop = 10, w = 1, h2 = 0.1, zstar = 1, o = 3))) %>%
  mutate(ic = list(c(10, 1))) %>% 
  mutate(sol = map2(pars, ic, ~ode(.y, seq(0, 100, by = 0.01), eqs, .x) %>%
                      as.data.frame() %>%
                      as_tibble() %>%
                      rename(N = `1`, mu = `2`) %>%
                      pivot_longer(cols = c(N, mu)) %>%
                      mutate(name = fct_relevel(name, "N", "mu")))) %>%
  unnest(sol) %>% 
  ggplot(aes(x = time, y = value)) +
  geom_line(colour = "steelblue") +
  facet_wrap(.~name, scales = "free_y")
```

### Starting with ideal cod body size, intense fishing ($\eta = 2$) and hypoxia ($\omicron = 0$):

```{r healthy_with_intense_fishing_hypoxia, fig.width = 6, out.width = "67%"}
tibble(pars = list(list(mF = 0, s = 1/2, theta = 5, eta = 2, alpha0 = 1 / sqrt(2), rho = 15, sF = 1/3, Fpop = 10, w = 1, h2 = 0.1, zstar = 1, o = 3))) %>%
  mutate(ic = list(c(10, 1))) %>% 
  mutate(sol = map2(pars, ic, ~ode(.y, seq(0, 100, by = 0.01), eqs, .x) %>%
                      as.data.frame() %>%
                      as_tibble() %>%
                      rename(N = `1`, mu = `2`) %>%
                      pivot_longer(cols = c(N, mu)) %>%
                      mutate(name = fct_relevel(name, "N", "mu")))) %>%
  unnest(sol) %>% 
  ggplot(aes(x = time, y = value)) +
  geom_line(colour = "steelblue") +
  facet_wrap(.~name, scales = "free_y")
```

## Starting with small cod body size ($\mu = -1$), no fishing pressure ($\eta = 0$) and hypoxia ($\omicron = 3$), with flounder competition:

```{r lowcondition_without_fishing_hypoxia, fig.width = 6, out.width = "67%"}
tibble(pars = list(list(mF = 0, s = 1/2, theta = 5, eta = 0, alpha0 = 1 / sqrt(2), rho = 15, sF = 1/3, Fpop = 10, w = 1, h2 = 0.1, zstar = 1, o = 3))) %>%
  mutate(ic = list(c(10, -1))) %>%
  mutate(sol = map2(pars, ic, ~ode(.y, seq(0, 100, by = 0.01), eqs, .x) %>%
                      as.data.frame() %>%
                      as_tibble() %>%
                      rename(N = `1`, mu = `2`) %>%
                      pivot_longer(cols = c(N, mu)) %>%
                      mutate(name = fct_relevel(name, "N", "mu")))) %>%
  unnest(sol) %>% 
  ggplot(aes(x = time, y = value)) +
  geom_line(colour = "steelblue") +
  facet_wrap(.~name, scales = "free_y")
```

### Starting with small cod body size ($\mu = -1$), no fishing, with hypoxia and reduced flounder competition:

```{r lowcondition_nofishing_lowflounder_hypoxia, fig.width = 6, out.width = "67%"}
tibble(pars = list(list(mF = 0, s = 1/2, theta = 5, eta = 0, alpha0 = 1 / sqrt(2), rho = 15, sF = 1/3, Fpop = 4, w = 1, h2 = 0.1, zstar = 1, o = 3))) %>%
  mutate(ic = list(c(15, -1))) %>%
  mutate(sol = map2(pars, ic, ~ode(.y, seq(0, 200, by = 0.01), eqs, .x) %>%
                      as.data.frame() %>%
                      as_tibble() %>%
                      rename(N = `1`, mu = `2`) %>%
                      pivot_longer(cols = c(N, mu)) %>%
                      mutate(name = fct_relevel(name, "N", "mu")))) %>%
  unnest(sol) %>%
  ggplot(aes(x = time, y = value)) +
  geom_line(colour = "steelblue") +
  facet_wrap(.~name, scales = "free_y")
```

